---
title: "Auditable qPCR Cleaning with tidyqpcr (v5: multi‑plate)"
format:
  html:
    toc: true
    code-fold: true
    css: styles.css
execute:
  echo: true
  warning: false
  message: false
params:
  # --- INPUTS ---
  input_dir: "data_raw"          # folder with plates
  file_pattern: "\\.csv$"       # regex for files to include
  files: null                     # optional explicit vector of files; overrides input_dir + pattern

  # --- OUTPUTS ---
  output_dir: "outputs"
  log_path: "audit/pcr_decisions.csv"

  # --- QC THRESHOLDS ---
  LOD_Lo: 0.012
  LOD_Hi: 2000
  delta_cq_threshold: 1.0

  # --- OPTIONAL REMOVAL BY PATTERN ---
  rm_pattern: ["OM"]        # e.g., ["OM", "^CTRL", "_test$"]
  rm_columns: ["sample"]    # e.g., ["sample","content"]
  enable_preview: false     # show preview tables for regex removal
  dry_run: false            # do NOT remove rows; only preview/log summary
  debug_print: false

  # --- REPORT LAYOUT ---
  pagebreak_each_plate: false
---

## Overview

This document processes **multiple plates** in a folder, treating **each file as an isolated run**. Results are **not combined**: every plate gets its **own section**, its **own output CSVs** named from the file stem, and all decisions are append‑logged with the **input file name** for traceability (MIQE‑aligned foundations via **tidyqpcr**).\
It also supports **regex‑based removal** with **preview** and **dry‑run** modes.

::: callout-tip
**Tip** — Configure which files to run in the YAML params: set `files:` explicitly, or use `input_dir` + `file_pattern`.
:::

```{=html}
<style>
.plate-section {margin-block: 1.25rem}

.page-break {break-after: page; page-break-after: always}
</style>
```

## Setup

```{r}
library(tidyverse)
library(readr)
library(lubridate)
library(janitor)
library(tidyqpcr)  # tidy, MIQE-aware workflow foundation
library(tools)
library(htmltools)
library(gt)
library(knitr)
```

### Helpers: logging (includes per-plate input file), reading, and processing

```{r}
# ---- Logging helpers ----
ensure_log_store <- function(path) {
  if (!dir.exists(dirname(path))) dir.create(dirname(path), recursive = TRUE)
  if (!file.exists(path)) {
    write_csv(
      tibble(
        timestamp  = as_datetime(character()),
        user       = Sys.info()[["user"]],
        run_id     = character(),
        input_file = character(),
        sample_id  = character(),
        target     = character(),
        rule_id    = character(),
        outcome    = character(),
        evidence   = character(),
        source     = "quarto"
      ),
      path
    )
  }
  invisible(path)
}

# The current input file name can change per-plate; pick it up from a global-like var if present
.resolve_input_file <- function(default_file) {
  val <- get0(".current_input_file", envir = .GlobalEnv, inherits = FALSE)
  if (is.null(val)) basename(default_file) else basename(val)
}

log_decision <- function(sample_id, target, rule_id, outcome, evidence,
                         run_id, log_path, default_file) {
  ensure_log_store(log_path)
  entry <- tibble(
    timestamp  = now(tzone = "UTC"),
    user       = Sys.info()[["user"]],
    run_id     = run_id,
    input_file = .resolve_input_file(default_file),
    sample_id  = as.character(sample_id %||% NA_character_),
    target     = as.character(target %||% NA_character_),
    rule_id    = rule_id,
    outcome    = outcome,
    evidence   = evidence,
    source     = "quarto"
  )
  write_csv(entry, log_path, append = TRUE)
  invisible(entry)
}

# ---- Core processing for a single plate file ----
process_plate <- function(file,
                          LOD_Lo, LOD_Hi, dCq_thr,
                          rm_patterns, rm_cols_req,
                          enable_preview, dry_run, debug_print,
                          output_dir, log_path) {
  assign(".current_input_file", file, envir = .GlobalEnv)

  stem   <- tools::file_path_sans_ext(basename(file))
  run_id <- paste0(format(Sys.time(), "%Y%m%d_%H%M%S"), "_", stem)

  # Import
  raw <- read_csv(file, show_col_types = FALSE)
  raw <- raw[19:nrow(raw), ]
  colnames(raw) <- raw[1, ]
  raw <- raw[2:nrow(raw), ]
  
  raw <- raw %>%
    janitor::clean_names()
  
  if(isTRUE(debug_print)) {
    print("Table: `raw`")
    print(head(raw))
  }
  
  stopifnot(all(c("well","fluor","target","content","sample","cq","starting_quantity_sq") %in% names(raw)))
  dat <- raw %>% mutate(
    cq_num = suppressWarnings(as.numeric(cq)),
    sq_raw = suppressWarnings(as.numeric(starting_quantity_sq))
  )
  
  if(isTRUE(debug_print)) {
    print("Table: `dat`")
    print(head(dat))
  }
  
  # Step 0: Removal of unnamed samples
  dat <- dat %>%
    mutate(
        sample_reason = case_when(
            is.na(sample) ~ "RM_Sample_removed_due_to_absent_name",
            TRUE ~ NA_character_
            ))
  
  dat %>%
    filter(!is.na(.data$sample_reason)) %>%
    rowwise() %>%
    do({
      log_decision(.$sample, .$target, .$sample_reason, "applied",
                   paste0("Removed due to absent sample name"),
                   run_id, log_path, file)
      tibble(.)
    }) %>% invisible()
  
  dat <- dat %>%
    filter(is.na(sample_reason))

  # Step 1: SQ adjust
  dat1 <- dat %>%
    mutate(
      sq_adj = case_when(
        is.na(sq_raw) | is.nan(sq_raw) ~ LOD_Lo/2,
        sq_raw < LOD_Lo                ~ LOD_Lo/2,
        TRUE                           ~ sq_raw
      ),
      sq_adj_reason = case_when(
        is.na(sq_raw) | is.nan(sq_raw) ~ "ADJ2:SQ_NA_to_half_LOD_Lo",
        sq_raw < LOD_Lo                ~ "ADJ1:SQ_below_LOD_Lo_to_half",
        TRUE                           ~ NA_character_
      )
    )
  
  #### TODO: Add in per specieis LOD (Lo and Hi) dict

  dat1 %>%
    filter(!is.na(.data$sq_adj_reason)) %>%
    rowwise() %>%
    do({
      log_decision(.$sample, .$target, .$sq_adj_reason, "applied",
                   paste0("sq_raw=", .$sq_raw, "; LOD_Lo=", LOD_Lo),
                   run_id, log_path, file)
      tibble(.)
    }) %>% invisible()
  
  if(isTRUE(debug_print)) {
    print("Table: `dat1 step1`")
    print(head(dat1))
  }

  # Step 2: Optional removal by regex pattern
  if (!is.null(rm_patterns) && length(rm_patterns) > 0) {
    rm_cols <- intersect(tolower(rm_cols_req), names(dat1))
    if (length(rm_cols) > 0) {
      combined_pat <- paste(rm_patterns, collapse = "|")
      re <- stringr::regex(combined_pat, ignore_case = TRUE)

      dat1 <- dat1 %>%
        mutate(rm_hit = if_any(all_of(rm_cols), ~ stringr::str_detect(as.character(.x), re)))

      if (isTRUE(enable_preview)) {
        rm_preview_counts <- dat1 %>% summarise(total_rows = n(), n_matches = sum(rm_hit, na.rm = TRUE),
                                                pct_matches = round(100 * n_matches / total_rows, 2))
        rm_counts_by_col <- map_dfr(rm_cols, function(cl) {
          tibble(column = cl,
                 n_matches = sum(stringr::str_detect(as.character(dat1[[cl]]), re), na.rm = TRUE))
        }) %>% arrange(desc(n_matches))
        rm_examples <- dat1 %>% filter(rm_hit) %>% select(well, fluor, target, content, sample, cq, starting_quantity_sq) %>% head(20)

        # Print preview tables to report
        cat("\n**Removal preview (", stem, ")**\n\n", sep = "")
        print(rm_preview_counts)
        print(rm_counts_by_col)
        print(rm_examples)

        # Log run-level preview
        log_decision(NA, NA, "RM_PATTERN", "preview",
                     paste0("patterns=", combined_pat, "; cols=", paste(rm_cols, collapse=","),
                            "; n_matches=", rm_preview_counts$n_matches),
                     run_id, log_path, file)
      }

      if (isTRUE(dry_run)) {
        n_hits <- sum(dat1$rm_hit, na.rm = TRUE)
        log_decision(NA, NA, "RM_PATTERN", "dry_run",
                     paste0("patterns=", combined_pat, "; cols=", paste(rm_cols, collapse=","),
                            "; n_matches=", n_hits),
                     run_id, log_path, file)
        dat1 <- dat1 %>% select(-rm_hit)
      } else {
        dat1 %>% filter(rm_hit) %>% rowwise() %>% do({
          log_decision(.$sample, .$target, "RM_PATTERN", "applied",
                       paste0("cols=", paste(rm_cols, collapse=","), "; pattern=", combined_pat),
                       run_id, log_path, file)
          tibble(.)
        }) %>% invisible()
        dat1 <- dat1 %>% filter(!rm_hit) %>% select(-rm_hit)
      }
    }
  }
  
  if(isTRUE(debug_print)) {
    print("Table: `dat1 step2`")
    print(head(dat1))
  }

  # Step 3: Sample naming consistency within (Fluor, Target, Content)
  name_checks <- dat1 %>%
    group_by(fluor, target, content) %>%
    summarise(n_samples = n_distinct(sample),
              samples   = paste(sort(unique(sample)), collapse = "|"), .groups = "drop") %>%
    mutate(flag_name_mismatch = n_samples > 1)

  name_checks %>% filter(flag_name_mismatch) %>% rowwise() %>% do({
    log_decision(NA, .$target, "QC_SN_MISMATCH", "applied",
                 paste0("content=", .$content, "; samples=", .$samples),
                 run_id, log_path, file)
    tibble(.)
  }) %>% invisible()

  dat2 <- dat1 %>%
    left_join(name_checks %>% select(fluor, target, content, flag_name_mismatch),
              by = c("fluor","target","content")) %>%
    mutate(flag_name_mismatch = replace_na(flag_name_mismatch, FALSE))
  
  if(isTRUE(debug_print)) {
    print("Table: `dat1 step3`")
    print(head(dat1))
  }
  
  if(isTRUE(debug_print)) {
    print("Table: `dat2 step3`")
    print(head(dat2))
  }

  # Steps 4 & 5: AverageSQ, DeltaCq, replicate health
  rep_key <- c("fluor","target","sample")

  rep_summary <- dat2 %>%
    group_by(across(all_of(rep_key))) %>%
    summarise(
      n_rows     = n(),
      n_num_cq   = sum(!is.na(cq_num) & !is.nan(cq_num)),
      cq_vals    = list(cq_num[!is.na(cq_num) & !is.nan(cq_num)]),
      delta_cq   = if (length(cq_vals[[1]]) == 2) abs(diff(range(cq_vals[[1]]))) else NA_real_,
      avg_sq     = mean(sq_adj, na.rm = TRUE),
      both_cq_na = (n_num_cq == 0),
      one_cq_na  = (n_num_cq == 1 & n_rows >= 2),
      single_rep = (n_rows == 1),
      .groups = "drop"
    ) %>%
    mutate(
      rv_single_rep   = single_rep,
      rv_delta_cq     = !is.na(delta_cq) & (delta_cq > dCq_thr) & (avg_sq > LOD_Lo),
      rv_mixed_na_num = one_cq_na,
      pass_negative   = both_cq_na,
      rv_high_sq      = avg_sq > LOD_Hi
    )

  rep_summary %>% rowwise() %>% do({
    if (.$rv_single_rep)
      log_decision(.$sample, .$target, "RV_SINGLE_REP", "applied",
                   paste0("n_rows=", .$n_rows), run_id, log_path, file)
    else
      log_decision(.$sample, .$target, "RV_SINGLE_REP", "skipped",
                   paste0("n_rows=", .$n_rows), run_id, log_path, file)

    if (!is.na(.$delta_cq) && (.$delta_cq > dCq_thr) && (.$avg_sq > LOD_Lo))
      log_decision(.$sample, .$target, "RV_DELTA_CQ", "applied",
                   paste0("|DeltaCq|=", .$delta_cq, "; thr=", dCq_thr,
                          "; avg_sq=", .$avg_sq, "; LOD_Lo=", LOD_Lo),
                   run_id, log_path, file)
    else
      log_decision(.$sample, .$target, "RV_DELTA_CQ", "skipped",
                   paste0("|DeltaCq|=", .$delta_cq, "; thr=", dCq_thr,
                          "; avg_sq=", .$avg_sq),
                   run_id, log_path, file)

    if (.$rv_mixed_na_num)
      log_decision(.$sample, .$target, "RV_MIXED_CQ", "applied",
                   paste0("n_num_cq=", .$n_num_cq, "; n_rows=", .$n_rows),
                   run_id, log_path, file)
    else
      log_decision(.$sample, .$target, "RV_MIXED_CQ", "skipped",
                   paste0("n_num_cq=", .$n_num_cq), run_id, log_path, file)

    if (.$pass_negative)
      log_decision(.$sample, .$target, "PASS_NEGATIVE", "pass",
                   "both Cq are NA/NaN", run_id, log_path, file)

    if (.$rv_high_sq)
      log_decision(.$sample, .$target, "RV_HIGH_SQ", "applied",
                   paste0("avg_sq=", .$avg_sq, "; LOD_Hi=", LOD_Hi),
                   run_id, log_path, file)
    else
      log_decision(.$sample, .$target, "RV_HIGH_SQ", "skipped",
                   paste0("avg_sq=", .$avg_sq), run_id, log_path, file)

    tibble(.)
  }) %>% invisible()
  
  if(isTRUE(debug_print)) {
    print("Table: `rep_summary`")
    print(head(rep_summary))
  }

  # Collate & present only on first replicate row
  dat3 <- dat2 %>%
    left_join(rep_summary %>%
                select(all_of(rep_key), delta_cq, avg_sq, both_cq_na,
                       rv_single_rep, rv_delta_cq, rv_mixed_na_num, rv_high_sq),
              by = rep_key) %>%
    mutate(flag_name_mismatch = coalesce(flag_name_mismatch, FALSE))

  reason_map <- tribble(
      ~flag,               ~reason,
    "flag_name_mismatch","Sample name mismatch within (Fluor, Target, Content)",
    "rv_single_rep",     "Only one replicate present",
    "rv_delta_cq",       "|DeltaCq| greater than threshold AND AverageSQ > LOD_Lo",
    "rv_mixed_na_num",   "One Cq is NA/NaN and the other numeric",
    "rv_high_sq",        "AverageSQ > LOD_Hi"
  )

  review_cols <- c("flag_name_mismatch","rv_single_rep","rv_delta_cq","rv_mixed_na_num","rv_high_sq")

  dat3 <- dat3 %>% mutate(
    needs_review = if_any(all_of(review_cols), ~ .x),
    review_reason = pmap_chr(
      pick(all_of(review_cols)),
      function(...) {
        flags_logical <- c(...)
        flags <- review_cols[flags_logical]
        if (length(flags) == 0) return(NA_character_)
        reasons <- reason_map$reason[match(flags, reason_map$flag)]
        paste(reasons, collapse = " ; ")
      }
    )
  )
  
  if(isTRUE(debug_print)) {
    print("Table: `dat3`")
    print(head(dat3))
  }

  dat4 <- dat3 %>%
    group_by(across(all_of(rep_key))) %>% arrange(well, .by_group = TRUE) %>%
    mutate(rep_idx = row_number()) %>% ungroup()
  
  if(isTRUE(debug_print)) {
    print("Table: `dat4`")
    print(head(dat4))
  }

  needs_review_grp <- dat4 %>%
    group_by(across(all_of(rep_key))) %>% summarise(needs_review_grp = any(needs_review), .groups = "drop")

  dat_out <- dat4 %>% mutate(
    out_delta_cq   = if_else(rep_idx == 1, delta_cq, NA_real_),
    out_average_sq = if_else(rep_idx == 1, avg_sq,   NA_real_),
    out_reason     = if_else(rep_idx == 1, review_reason, NA_character_)
  ) %>% left_join(needs_review_grp, by = rep_key)
  
  if(isTRUE(debug_print)) {
    print("Table: `dat_out`")
    print(head(dat_out))
  }

  dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
  all_path    <- file.path(output_dir, paste0(stem, "_all_samples.csv"))
  review_path <- file.path(output_dir, paste0(stem, "_review_samples.csv"))

  all_samples <- dat_out %>% transmute(
    Well   = well,
    Fluor  = fluor,
    Target = target,
    Content= content,
    Sample = sample,
    Cq     = cq,
    `Starting Quantity (SQ)` = sq_adj,
    DeltaCq   = out_delta_cq,
    AverageSQ = out_average_sq,
    review_reason = out_reason
  )
  
  if(isTRUE(debug_print)) {
    print("Table: `all_samples`")
    print(head(all_samples))
  }

  review_samples <- all_samples %>%
    left_join(dat_out %>% select(well, fluor, target, sample, needs_review_grp),
              by = c("Well"="well","Fluor"="fluor","Target"="target","Sample"="sample")) %>%
    filter(needs_review_grp) %>% select(-needs_review_grp)

  if(isTRUE(debug_print)) {
    print("Table: `review_samples`")
    print(head(review_samples))
  }

  write_csv(all_samples, all_path)
  write_csv(review_samples, review_path)

  list(
    stem = stem,
    all_path = all_path,
    review_path = review_path,
    n_in = nrow(raw),
    n_out = nrow(all_samples),
    n_review = nrow(review_samples),
    preview_enabled = isTRUE(enable_preview),
    dry_run = isTRUE(dry_run)
  )
}
```

### Discover plate files to process

```{r}
plate_files <- if (!is.null(params$files)) {
  # explicit vector supplied
  as.character(params$files)
} else {
  list.files(params$input_dir, pattern = params$file_pattern, full.names = TRUE)
}

validate <- file.exists(plate_files)
if (length(plate_files) == 0L || !all(validate)) {
  stop("No input files found (or some files do not exist). Check params$files / params$input_dir / params$file_pattern.")
}

plate_files
```

## Per-plate processing

```{r}
#| echo: false
#| message: false
#| warning: false
#| results: asis

render_plate_section <- function(stem, res) {
  summary_tbl <- tibble(
    File                 = basename(res$file),
    `Rows (input)`       = res$n_in,
    `Rows (output)`      = res$n_out,
    `Rows (review)`      = res$n_review,
    `All samples CSV`    = res$all_path,
    `Review samples CSV` = res$review_path
  )

  gt_tbl <- summary_tbl %>%
    gt() %>%
    tab_header(
      title = md("**Plate summary**"),
      subtitle = md(paste0("*", stem, "*"))
    ) %>%
    cols_align(align = "left", columns = everything()) %>%
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_column_labels(everything())
    ) %>%
    fmt_integer(columns = c(`Rows (input)`, `Rows (output)`, `Rows (review)`)) %>%
    opt_table_outline() %>%
    tab_options(table.font.size = px(14), data_row.padding = px(6))

  # Start fenced div with a class for styling/layout if desired
  knitr::asis_output('::: {.plate-section}\n')

  # Heading: use markdown so it’s portable across HTML and PDF
  knitr::asis_output(paste0("## Plate: ", stem, "\n\n"))

  # Print the gt table (in asis mode it will emit HTML/LaTeX appropriately)
  print(gt_tbl)

  # Close fenced div
  knitr::asis_output("\n:::\n")
}

for (i in seq_along(plate_files)) {
  f    <- plate_files[[i]]
  stem <- tools::file_path_sans_ext(basename(f))

  res <- process_plate(
    file = f,
    LOD_Lo = params$LOD_Lo, LOD_Hi = params$LOD_Hi, dCq_thr = params$delta_cq_threshold,
    rm_patterns = params$rm_pattern, rm_cols_req = params$rm_columns,
    enable_preview = params$enable_preview, dry_run = params$dry_run,
    debug_print = params$debug_print, output_dir = params$output_dir,
    log_path = params$log_path
  )
  # Keep the file path for the table
  res$file <- f

  render_plate_section(stem, res)

  # Optional page break between sections
  if (isTRUE(params$pagebreak_each_plate) && i < length(plate_files)) {
    if (knitr::is_html_output()) {
      knitr::asis_output("\n<div class='page-break'></div>\n")
    } else if (knitr::is_latex_output()) {
      knitr::asis_output("\n\\newpage\n")
    }
  }
}
```

## Appendix — Latest audit entries & counts (all plates in this run)

```{r}
audit <- read_csv(params$log_path, show_col_types = FALSE)
audit %>% arrange(desc(timestamp)) %>% head(30)

audit %>% count(input_file, rule_id, outcome, sort = TRUE)
```
